The evolution of gene expression in the term placenta of viviparous mammals

### 2016-05-16 manuscript/manuscript.Rnw
don't include figures
fix double include typo
start working on reformatting for genes & development

%\documentclass[english,12pt]{article}
\documentclass[doublespacing]{bmcart_dla}
\usepackage{fontspec}
\setmainfont{FreeSans}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSans}
\setmathsf{FreeSans}
\setmathtt{FreeSans}
%\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=14pt]{geometry}
\usepackage[letterpaper]{geometry}
\usepackage{url}
% \usepackage{fancyhdr}
\usepackage{graphicx}
% \usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
%\usepackage[plain]{fancyref}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage{sansmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=numeric-comp,style=nature,autocite=inline]{biblatex}
\addbibresource{references.bib}
% \usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
\usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
\renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
\usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}
% \newenvironment{narrow}[2]{%
%   \begin{list}{}{%
%       \setlength{\topsep}{0pt}%
%       \setlength{\leftmargin}{#1}%
%       \setlength{\rightmargin}{#2}%
%       \setlength{\listparindent}{\parindent}%
%       \setlength{\itemindent}{\parindent}%
%       \setlength{\parsep}{\parskip}}%
%   \item[]}{\end{list}}
% \def\newblock{\hskip}
% \newenvironment{paperquote}{%
%   \begin{quote}%
%      \it
%   }%
%   {\end{quote}}
% \renewcommand{\textfraction}{0.15}
% \renewcommand{\topfraction}{0.85}
% \renewcommand{\bottomfraction}{0.65}
% \renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
% \newenvironment{enumerate*}%
%   {\begin{enumerate}%
%     \setlength{\itemsep}{0pt}%
%     \setlength{\parskip}{0pt}}%
%   {\end{enumerate}}
% \newenvironment{itemize*}%
%   {\begin{itemize}%
%     \setlength{\itemsep}{0pt}%
%     \setlength{\parskip}{0pt}}%
%   {\end{itemize}}
% \setcounter{tocdepth}{3}
% \oddsidemargin 0.0in 
% \textwidth 6in
% \raggedbottom
% \clubpenalty = 10000
% \widowpenalty = 10000
% set a large emergency stretch to avoid issues with overfull boxes
% \setlength{\emergencystretch}{20em}
% make the afill smaller
% \renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
% \pagestyle{fancy}

@@

\begin{document}
\raggedright

@@

\begin{frontmatter}
\begin{fmbox}
\dochead{Research}%
\title{The evolution of gene expression in the term placenta of viviparous mammals}%
%
\author[addressref={aff1},noteref={n1},email={don@donarmstrong.com}]{\inits{DL}\fnm{Don L} \snm{Armstrong}}
\author[addressref={aff2},noteref={n1},email={1mmcgowen1@gmail.com}]{\inits{MR}\fnm{Michael R.} \snm{McGowen}}
\author[addressref={aff1},email={aweckle@illinois.edu}]{\inits{AL}\fnm{Amy} \snm{Weckle}}
\author[addressref={aff3},email={jacaravas@gmail.com}]{\inits{J}\fnm{Jason} \snm{Caravas}}
\author[addressref={aff4},email={agnewd@dcpah.msu.edu}]{\inits{D}\fnm{Dalen} \snm{Agnew}}
\author[addressref={aff5},email={kbenirschke@ucsd.edu}]{\inits{K}\fnm{Kurt} \snm{Benirschke}}
\author[addressref={aff6},email={suerumbaugh@gmail.com}]{\inits{S}\fnm{Sue} \snm{Savage-Rumbaugh}}
\author[addressref={aff11},email={nevo@research.haifa.ac.il}]{\inits{E}\fnm{Eviatar} \snm{Nevo}}
\author[addressref={aff7,aff8},email={cjkim@med.wayne.edu}]{\inits{CJ}\fnm{Chong J} \snm{Kim}}
\author[addressref={aff9},email={gunter.wagner@yale.edu}]{\inits{GP}\fnm{Günter P.} \snm{Wagner}}
\author[addressref={aff7},email={prbchiefstaff@med.wayne.edu}]{\inits{R}\fnm{Roberto} \snm{Romero}}
\author[addressref={aff1,aff6,aff10},corref={aff1},email={wildmand@illinois.edu}]{\inits{DE}\fnm{Derek E.} \snm{Wildman}}
\address[id=aff1]{Institute for Genomic Biology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
\address[id=aff2]{School of Biological and Chemical Sciences, Queen Mary, University of London, London, UK}
\address[id=aff3]{Center for Molecular Medicine and Genetics, Wayne State University School of Medicine, Detroit, MI, USA}
\address[id=aff4]{Diagnostic Center for Population and Animal Health, College of Veterinary Medicine, Michigan State University, East Lansing, MI, USA}
\address[id=aff5]{Department of Pathology, University of California San Diego, La Jolla, CA, USA}
\address[id=aff6]{Great Ape Trust/Bonobo Hope Sanctuary, Des Moines, IA, USA}
\address[id=aff7]{Perinatology Research Branch, Eunice Kennedy Shriver National Institute of Child Health and Human Development (NICHD), National Institutes of Health (NIH), Detroit, MI, USA}
\address[id=aff8]{Department of Pathology, Wayne State University School of Medicine, Detroit, MI, USA}
\address[id=aff9]{Department of Ecology and Evolutionary Biology and Yale Systems Biology Institute, Yale University, New Haven, CT, USA}
\address[id=aff10]{Department of Molecular and Integrative Physiology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
\address[id=aff11]{The Institute of Evolution, University of Haifa, Haifa, Israel}

\begin{artnotes}
\end{artnotes}

\begin{artnotes}
\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}
% \twocolumn

\begin{abstractbox}
\begin{abstract}
% \section{Abstract}
  \parttitle{Background} The placenta is arguably the most
  anatomically variable organ in mammals. Understanding gene
  expression patterns among placentas from multiple species
  illuminates the molecular basis of placenta function as well
  obstetrical syndromes affecting humans.

  \parttitle{Results} We measured the expression profiles of
  \Sexpr{star.logs[,length(unique(file))]} term placentas of
  \Sexpr{star.logs[,length(unique(species))]} species of mammals. We
  found lineage-specific differential expression of prolactin,
  relaxin, and galectins as well as increased expression of the
  pre-eclampsia associated genes \textit{ADAM12}, \textit{CRH}, and
  \textit{PAPPA} in humans. We also identified genes which are
  expressed in all placentas, including genes with immune-modulating
  properties (\textit{ANXA2}, \textit{ANXA1}, \textit{S100A11}, and
  \textit{S100A10}) and cell-cell interactions (\textit{LAMC1},
  \textit{LUM}, and \textit{LGALS1}).

  \parttitle{Conclusions} A diverse array of genes are differentially
  expressed and are likely to play a role in the anatomical
  variability of the placenta in the placentas of mammalian lineages.
  Furthermore, multiple genes which are associated with pre-eclampsia
  (\textit{CRH}, \textit{ADAM12}, \textit{KISS1}, \textit{PAPPA},
  \textit{IL1RL1}, and \textit{PLAC1}) show increased expression on
  the human lineage. Still, using the strictest filtering strategy
  we find a core set of only
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes is essential for the proper functioning of the placenta in all
  species analyzed.

@@

\end{abstract}
\end{abstractbox}
\end{frontmatter}

@@

    }
    return(invisible(temp.table[order(`p (t-test)`),]))
}
output.top.genes.table <- function(data,output=TRUE) {
    if (output) {
        table.display <- c("s","s","s","d")
        table.align <- c("c","c","p{3in}","r")
        if (any(colnames(data) %in% "SEM"))  {
            table.display <- c(table.display,"d")
            table.align <- c(table.align,"r")
        }
        if (any(colnames(data) %in% "Human Symbol"))  {
            table.align <- c("c",table.align)
            table.display <- c("s",table.display)
        }
        print(xtable(data,
                     display=table.display,
                     align=table.align
                     ),
              booktabs=TRUE,
              floating=FALSE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)
    }
}

@@

\begin{table}
  \footnotesize
<<top.50.human.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="homo sapiens"))
@ 
\caption{The top 50 genes by \ac{FPKM} in \human which are not
  annotated as Mt\_rRNA, Mt\_tRNA, rRNA or a housekeeping gene. Gene
  symbol is the gene symbol or Ensembl gene id. Human Name is the HUGO
  name corresponding to the human symbol. A complete list of all genes
  is in \cref{tab:all_homo_sapiens_placenta}.}
\label{tab:top50_human_placenta}
\end{table}

@@

\begin{table}
  \footnotesize
<<top.50.bonobo.expressed.genes,echo=FALSE,results="asis">>=
top.bonobo.genes <- top.genes.table(species.name="pan paniscus")
output.top.genes.table(top.bonobo.genes)
setkey(top.bonobo.genes,"Gene symbol")
@ 
\caption{\tableexplanation{\bonobo}{tab:all_pan_paniscus_placenta}}
\label{tab:top50_bonobo_placenta}
\end{table}

@@

\begin{table}
  \footnotesize
<<top.50.spider.monkey.expressed.genes,echo=FALSE,results="asis">>=
top.spider.monkey.genes <- top.genes.table(species.name="ateles fusciceps")
output.top.genes.table(top.spider.monkey.genes)
setkey(top.spider.monkey.genes,"Gene symbol")
@ 
\caption{\tableexplanation{\smonkey}{tab:all_ateles_fusciceps_placenta}}
\label{tab:top50_spider_monkey_placenta}
\end{table}

@@

\begin{table}[h]
  \footnotesize
<<top.50.mouse.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="mus musculus"))
@ 
\caption{\tableexplanation{\mouse}{tab:all_mus_musculus_placenta}}
\label{tab:top50_mouse_placenta}
\end{table}

@@

\begin{table}[h]
  \footnotesize
<<top.50.n.galili.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="spalax galili"))
@ 
\caption{\tableexplanation{\galili}{tab:all_spalax_galili_placenta}}
\label{tab:top50_n_galili_placenta}
\end{table}

@@

\begin{table}[h]
  \footnotesize
<<top.50.s.carmeli.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="spalax carmeli"))
@ 
\caption{\tableexplanation{\carmeli}{tab:all_spalax_carmeli_placenta}}
\label{tab:top50_s_carmeli_placenta}
\end{table}

@@

\begin{table}
  \footnotesize
<<top.50.dog.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="canis familiaris"))
@ 
\caption{\tableexplanation{\dog}{tab:all_canis_familiaris_placenta}}
\label{tab:top50_dog_placenta}
\end{table}

@@

\begin{table}
  \footnotesize
<<top.50.cow.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="bos taurus"))
@ 
\caption{\tableexplanation{\cow}{tab:all_bos_taurus_placenta}}
\label{tab:top50_cow_placenta}
\end{table}

@@

\begin{table}[h]
  \footnotesize
<<top.50.pig.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="sus scrofa"))
@ 
\caption{\tableexplanation{\pig}{tab:all_sus_scrofa_placenta}}
\label{tab:top50_pig_placenta}
\end{table}

@@

\begin{table}[h]
  \footnotesize
<<top.50.horse.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="equus caballus"))
@ 
\caption{\tableexplanation{\horse}{tab:all_equus_caballus_placenta}}
\label{tab:top50_horse_placenta}
\end{table}

@@

\begin{table}[h]
  \footnotesize
<<top.50.sheep.expressed.genes,echo=FALSE,results="asis">>=
top.sheep.genes <- top.genes.table(species.name="ovis aries",n=200)
top.sheep.genes <-
    top.sheep.genes[!grepl("\\{(7SK|.+_([SM]RP|nuc)|SNOR[AD]\\d+|U\\d+|HIST.*)\\}",
                           `Gene symbol`) &
                    !grepl("\\((7SK|.+_([SM]RP|nuc)|SNOR[AD]\\d+|U\\d+|HIST.*)\\)",
                           `Human Symbol`)
                   ,][1:50,]
output.top.genes.table(top.sheep.genes)
@ 
\caption{\tableexplanation{\sheep}{tab:all_ovis_aries_placenta}}
\label{tab:top50_sheep_placenta}
\end{table}

@@

\begin{table}
  \footnotesize
<<top.50.opossum.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="monodelphis domestica"))
@ 
\caption{\tableexplanation{\opossum}{tab:all_monodelphis_domestica_placenta}}
\label{tab:top50_opossum_placenta}
\end{table}

@@

The transcriptomes from multiple different species enabled us to
identify the set of expressed genes which are common to all placentas.
Components of the annexin complexes, including including
\textit{ANXA2}, \textit{ANXA1}, \textit{S100A11}, and \textit{S100A10}
are expressed in all placentas examined, and argue for the evolution
importance of annexins in placenta function, where they likely
establish maternal-fetal
tolerance\cite{Liu.ea2015:AnnexinA2complexeswith,Hutchinson.ea2011:Molecularregulatorsofresolution,Aarli.ea1997:Suppressiveeffectonlymphoproliferation}.
We suggest that further research into molecular role of these annexin
complexes may illuminate how immune tolerance is maintained during
placentation. The core placenta transcriptome is also enriched in
genes involved in maintaining the structure of the placenta and
transporting proteins, as would be expected given the function of the
placenta.

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\setmathrm{FreeSerif}
\setmathsf{FreeSans}
\setmathtt{FreeSans}
\usepackage[letterpaper,left=0.75in,right=0.75in,top=0.75in,bottom=0.75in,headheight=14pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
% \usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
%\usepackage[plain]{fancyref}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{amsmath}
\usepackage{sansmath}
\usepackage[backend=biber,natbib=true,hyperref=true,citestyle=authoryear-comp,firstinits=true,sorting=nyt,bibstyle=authoryear,autocite=inline,maxcitenames=1,mincitenames=1,maxbibnames=10,minbibnames=10]{biblatex}
\DefineBibliographyStrings{english}{andothers = {\emph{et\addabbrvspace al\adddot}}}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=black,citecolor=blue]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
\usepackage[nolists,figuresfirst,nomarkers,heads]{endfloat}
\renewcommand{\efloatseparator}{\mbox{}}
%\AtBeginFigures{\newpage \renewcommand{\efloatseparator}{\mbox{}}\section*{Figure Legends}}
%\AtBeginTables{\newpage \section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\renewcommand{\includegraphics}[2][]{}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\newcommand{\DEW}[1]{\textcolor{purple}{\fxnote{DEW: #1}}}
% \usepackage[x11names,svgnames]{xcolor}
% \usepackage{texshade}
% \usepackage{tikz}
\usepackage{nameref}
\usepackage{xspace}
% this package will stretch the font slightly if necessary.
\usepackage{microtype}
\usepackage{zref-xr,zref-user}
\zxrsetup{toltxlabel}
\zexternaldocument*{supplemental_material}
\usepackage[capitalize]{cleveref}
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% set a large emergency stretch to avoid issues with overfull boxes
% \setlength{\emergencystretch}{20em}
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
% \pagestyle{fancy}

@@

\begin{document}

@@

\title{The evolution of gene expression in the term placenta of viviparous mammals}%
%
%\author[addressref={aff1},noteref={n1},email={don@donarmstrong.com}]{\inits{DL}\fnm{Don L} \snm{Armstrong}}
\author[1,†]{Don L. Armstrong}
%\author[addressref={aff2},noteref={n1},email={1mmcgowen1@gmail.com}]{\inits{MR}\fnm{Michael R.} \snm{McGowen}}
\author[2,†]{Michael R. McGowen}
%\author[addressref={aff1},email={aweckle@illinois.edu}]{\inits{AL}\fnm{Amy} \snm{Weckle}}
\author[1]{Amy Weckle}
%\author[addressref={aff3},email={jacaravas@gmail.com}]{\inits{J}\fnm{Jason} \snm{Caravas}}
\author[3]{Jason Caravas}
%\author[addressref={aff4},email={agnewd@dcpah.msu.edu}]{\inits{D}\fnm{Dalen} \snm{Agnew}}
\author[4]{Dalen Agnew}
%\author[addressref={aff5},email={kbenirschke@ucsd.edu}]{\inits{K}\fnm{Kurt} \snm{Benirschke}}
\author[5]{Kurt Benirschke}
%\author[addressref={aff6},email={suerumbaugh@gmail.com}]{\inits{S}\fnm{Sue} \snm{Savage-Rumbaugh}}
\author[6]{Sue Savage-Rumbaugh}
%\author[addressref={aff11},email={nevo@research.haifa.ac.il}]{\inits{E}\fnm{Eviatar} \snm{Nevo}}
\author[7]{Eviatar Nevo}
%\author[addressref={aff7,aff8},email={cjkim@med.wayne.edu}]{\inits{CJ}\fnm{Chong J} \snm{Kim}}
\author[8,9]{Chong J. Kim}
%\author[addressref={aff9},email={gunter.wagner@yale.edu}]{\inits{GP}\fnm{Günter P.} \snm{Wagner}}
\author[10]{Gunter Wagner}
% \author[addressref={aff7},email={prbchiefstaff@med.wayne.edu}]{\inits{R}\fnm{Roberto} \snm{Romero}}
\author[8]{Roberto Romero}
%\author[addressref={aff1,aff6,aff10},corref={aff1},email={wildmand@illinois.edu}]{\inits{DE}\fnm{Derek E.} \snm{Wildman}}
\author[1,11,*]{Derek E. Wildman}
\affil[1]{Institute for Genomic Biology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
\affil[2]{School of Biological and Chemical Sciences, Queen Mary, University of London, London, UK}
\affil[3]{Center for Molecular Medicine and Genetics, Wayne State University School of Medicine, Detroit, MI, USA}
\affil[4]{Diagnostic Center for Population and Animal Health, College of Veterinary Medicine, Michigan State University, East Lansing, MI, USA}
\affil[5]{Department of Pathology, University of California San Diego, La Jolla, CA, USA}
\affil[6]{Great Ape Trust/Bonobo Hope Sanctuary, Des Moines, IA, USA}
\affil[7]{The Institute of Evolution, University of Haifa, Haifa, Israel}
\affil[8]{Perinatology Research Branch, Eunice Kennedy Shriver National Institute of Child Health and Human Development (NICHD), National Institutes of Health (NIH), Detroit, MI, USA}
\affil[9]{Department of Pathology, Wayne State University School of Medicine, Detroit, MI, USA}
\affil[10]{Department of Ecology and Evolutionary Biology and Yale Systems Biology Institute, Yale University, New Haven, CT, USA}
\affil[11]{Department of Molecular and Integrative Physiology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
\affil[†]{Equal contributors}
\affil[*]{To whom correspondence should be addressed: wildman@illinois.edu}

\maketitle

 \section{Abstract}
  The placenta is arguably the most
  anatomically variable organ in mammals. Understanding gene
  expression patterns among placentas from multiple species
  illuminates the molecular basis of placenta function as well
  obstetrical syndromes affecting humans.

  We measured the expression profiles of
  \Sexpr{star.logs[,length(unique(file))]} term placentas of
  \Sexpr{star.logs[,length(unique(species))]} species of mammals. We
  found lineage-specific differential expression of prolactin,
  relaxin, and galectins as well as increased expression of the
  pre-eclampsia associated genes \textit{ADAM12}, \textit{CRH}, and
  \textit{PAPPA} in humans. We also identified genes which are
  expressed in all placentas, including genes with immune-modulating
  properties (\textit{ANXA2}, \textit{ANXA1}, \textit{S100A11}, and
  \textit{S100A10}) and cell-cell interactions (\textit{LAMC1},
  \textit{LUM}, and \textit{LGALS1}).

  A diverse array of genes are differentially
  expressed and are likely to play a role in the anatomical
  variability of the placenta in the placentas of mammalian lineages.
  Furthermore, multiple genes which are associated with pre-eclampsia
  (\textit{CRH}, \textit{ADAM12}, \textit{KISS1}, \textit{PAPPA},
  \textit{IL1RL1}, and \textit{PLAC1}) show increased expression on
  the human lineage. Still, using the strictest filtering strategy
  we find a core set of only
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes is essential for the proper functioning of the placenta in all
  species analyzed.

@@

@@

    }
    return(invisible(temp.table[order(`p (t-test)`),]))
}
output.top.genes.table <- function(data,output=TRUE) {
    if (output) {
        table.display <- c("s","s","s","d")
        table.align <- c("c","c","p{4in}","r")
        if (any(colnames(data) %in% "SEM"))  {
            table.display <- c(table.display,"d")
            table.align <- c(table.align,"r")
        }
        if (any(colnames(data) %in% "Human Symbol"))  {
            table.align <- c("c",table.align)
            table.display <- c("s",table.display)
        }
        print(xtable(data,
                     display=table.display,
                     align=table.align
                     ),
              booktabs=TRUE,
              floating=FALSE,
              sanitize.text.function=genes.table.sanitize,
              include.rownames=FALSE)
    }
}

@@

\begin{table}
  \small
<<top.50.human.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="homo sapiens"))
@ 
\caption{The top 50 genes by \ac{FPKM} in \human which are not
  annotated as Mt\_rRNA, Mt\_tRNA, rRNA or a housekeeping gene. Gene
  symbol is the gene symbol or Ensembl gene id. Human Name is the HUGO
  name corresponding to the human symbol. A complete list of all genes
  is in \cref{tab:all_homo_sapiens_placenta}.}
\label{tab:top50_human_placenta}
\end{table}

@@

\begin{table}
  \small
<<top.50.bonobo.expressed.genes,echo=FALSE,results="asis">>=
top.bonobo.genes <- top.genes.table(species.name="pan paniscus")
output.top.genes.table(top.bonobo.genes)
setkey(top.bonobo.genes,"Gene symbol")
@ 
\caption{\tableexplanation{\bonobo}{tab:all_pan_paniscus_placenta}}
\label{tab:top50_bonobo_placenta}
\end{table}

@@

\begin{table}
  \small
<<top.50.spider.monkey.expressed.genes,echo=FALSE,results="asis">>=
top.spider.monkey.genes <- top.genes.table(species.name="ateles fusciceps")
output.top.genes.table(top.spider.monkey.genes)
setkey(top.spider.monkey.genes,"Gene symbol")
@ 
\caption{\tableexplanation{\smonkey}{tab:all_ateles_fusciceps_placenta}}
\label{tab:top50_spider_monkey_placenta}
\end{table}

@@

\begin{table}[h]
  \small
<<top.50.mouse.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="mus musculus"))
@ 
\caption{\tableexplanation{\mouse}{tab:all_mus_musculus_placenta}}
\label{tab:top50_mouse_placenta}
\end{table}

@@

\begin{table}[h]
  \small
<<top.50.n.galili.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="spalax galili"))
@ 
\caption{\tableexplanation{\galili}{tab:all_spalax_galili_placenta}}
\label{tab:top50_n_galili_placenta}
\end{table}

@@

\begin{table}[h]
  \small
<<top.50.s.carmeli.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="spalax carmeli"))
@ 
\caption{\tableexplanation{\carmeli}{tab:all_spalax_carmeli_placenta}}
\label{tab:top50_s_carmeli_placenta}
\end{table}

@@

\begin{table}
  \small
<<top.50.dog.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="canis familiaris"))
@ 
\caption{\tableexplanation{\dog}{tab:all_canis_familiaris_placenta}}
\label{tab:top50_dog_placenta}
\end{table}

@@

\begin{table}
  \small
<<top.50.cow.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="bos taurus"))
@ 
\caption{\tableexplanation{\cow}{tab:all_bos_taurus_placenta}}
\label{tab:top50_cow_placenta}
\end{table}

@@

\begin{table}[h]
  \small
<<top.50.pig.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="sus scrofa"))
@ 
\caption{\tableexplanation{\pig}{tab:all_sus_scrofa_placenta}}
\label{tab:top50_pig_placenta}
\end{table}

@@

\begin{table}[h]
  \small
<<top.50.horse.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="equus caballus"))
@ 
\caption{\tableexplanation{\horse}{tab:all_equus_caballus_placenta}}
\label{tab:top50_horse_placenta}
\end{table}

@@

\begin{table}[h]
  \small
<<top.50.sheep.expressed.genes,echo=FALSE,results="asis">>=
top.sheep.genes <- top.genes.table(species.name="ovis aries",n=200)
top.sheep.genes <-
    top.sheep.genes[!grepl("\\{(7SK|.+_([SM]RP|nuc)|SNOR[AD]\\d+|U\\d+|HIST.*)\\}",
                           `Gene symbol`) &
                    !grepl("\\((7SK|.+_([SM]RP|nuc)|SNOR[AD]\\d+|U\\d+|HIST.*)\\)",
                           `Human Symbol`)
                   ,][1:50,]
output.top.genes.table(top.sheep.genes)
@ 
\caption{\tableexplanation{\sheep}{tab:all_ovis_aries_placenta}}
\label{tab:top50_sheep_placenta}
\end{table}

@@

\begin{table}
  \small
<<top.50.opossum.expressed.genes,echo=FALSE,results="asis">>=
output.top.genes.table(top.genes.table(species.name="monodelphis domestica"))
@ 
\caption{\tableexplanation{\opossum}{tab:all_monodelphis_domestica_placenta}}
\label{tab:top50_opossum_placenta}
\end{table}

@@

The transcriptomes from multiple different species enabled us to
identify the set of expressed genes which are common to all placentas.
Components of the annexin complexes, including
\textit{ANXA2}, \textit{ANXA1}, \textit{S100A11}, and \textit{S100A10}
are expressed in all placentas examined, and argue for the evolution
importance of annexins in placenta function, where they likely
establish maternal-fetal
tolerance\cite{Liu.ea2015:AnnexinA2complexeswith,Hutchinson.ea2011:Molecularregulatorsofresolution,Aarli.ea1997:Suppressiveeffectonlymphoproliferation}.
We suggest that further research into molecular role of these annexin
complexes may illuminate how immune tolerance is maintained during
placentation. The core placenta transcriptome is also enriched in
genes involved in maintaining the structure of the placenta and
transporting proteins, as would be expected given the function of the
placenta.
